---
title: "Day 4"
author: "Derrick Theophille"
format: html
---

# Standardized CPUE

## Relative abundance indices

Catch data, nominal CPUE, and standardized CPUE are used as indices of abundance. Need at least 10 years for any reliable assessment.

CPUE can be:
- total catch per number of fishers (divers) per day; used for conch or lobsters
- total catch per number of fishing operations, sets, or hauls; used for operations such as fish pots or nets
- total catch per number of gear (hooks, pots, etc.); used for hook and line fisheries, but also pots
- total catch per number of fishing trips
- total catch per total duration of sets; used for pots or nets
- total catch 

## CPUE equation

Catch = qEN, where
q = the extent to which the stock is susceptible to fishing, AKA catchability
E = effort
N = abundance


CPUE = catch/E

## What is standardised CPUE?

This is the removal of the unnecessary effects on catchability inherent in things such as season, area (location), and gear. Nominal CPUE usually contains these unnecessary effects on catchability.

Generalized Linear Model (GLM) is used to determine standardised CPUE.

```{r}
#| label: packages

library(tidyverse)
library(here)
```

```{r}
#| label: calculate-standardiseed-CPUE

# load data
catch <- read.csv(here("00_raw_data", "nominalCPUE.csv"))

# estimate-standardised-CPUE
catch$nominal[catch$nominal <0] <- 0

plot(tapply(catch$nominal, catch$year, mean), type = "l",
     xlab = "Year",
     ylab = "Relative abundance",
     axes = F)
# par(new = T)
# plot(N, type = "l", lwd = 3, col = "lightgrey") # This line should show the estimated stock abundance

op <- glm(nominal ~ year + month + area, data = catch)
summary(op)

newdata <- expand.grid(year = c(2013:2022), 
                       month = c(1:12), 
                       area = c(1, 2))

# Create a new column to estimate catch amount for each row 
newdata$est <- predict(op, newdata)

par(new = T)
plot(tapply(newdata$est, newdata$year, mean), 
     type = "l", col = "blue",
     ylab = "",
     xlab = "")
```

```{r}
#| label: exercise

# load-data
catch_02 <- read.csv(here("00_raw_data", "nominalCPUE2.csv"))

# estimate standardised CPUE
catch_02 <- catch_02 |> 
  mutate(nominal = ifelse(nominal <0, 0, nominal))

op <- glm(nominal ~ year + month + area, data = catch_02)
summary(op)

# plot annual nominal CPUE and standardised  CPUE
plot(tapply(catch_02
            $nominal, catch_02$year, mean), 
     type = "l",
     xlab = "Year",
     ylab = "Relative abundance",
     axes = F)

summary(catch_02)

newdata <- expand.grid(year = c(2013:2022), 
                       month = c(1:12), 
                       area = c(1, 2))

# Create a new column to estimate catch amount for each row 
newdata$est <- predict(op, newdata)

# PLot
par(new = T)
plot(tapply(newdata$est, newdata$year, mean), 
     type = "l", col = "blue",
     ylab = "",
     xlab = "",
     axes = F)
axis(side = 1, at = c(1:10), c(2013:2022))
axis(side = 2, labels = F)
legend("topleft", legend = c("Nominal CPUE", "Standardized CPUE"), col = c("black", "blue"), lwd = c(2,2))


```


# CMSY

Catch Maximum Sustainable Yield (CMSY)



